### 2021-01-25 refereeResponse.txt
Move figures backward in text a little

We appreciate the referee's thorough and careful comments. We will address concerns with the manuscript one by one, with the original comment listed first, followed by our response (indented).

1) It is not clear how collisions were treated. Wallace et al. point to other papers that describe the implementations of the collisional algorithms. Although I agree they should not attempt to describe all details of the code here, it would be of major help to add more text describing how they deal with collisions.

	A more detailed discussion of the collision model has been added to the end of section 3.1. Most importantly, it is now explicitly stated that we only model the planetesimal-planetesimal collisions.

1.1) What is the collisional scaling used?

	All collisions result in perferct mergers (although the subsequent growth of the planetesimals does not matter much, as no bodies in any of the simulations happen to undergo more than a single collision). This is now stated in section 3.1. We do not model the subsequent evolution of any of the debris.

1.2) What material is assumed for the colliding planetesimals?

	A 2 g cm^-3 bulk density corresponds to an even mixture of ice and rock. Because of our simple collision model, this only affects the mass-radius relation of the bodies.

1.3) How do you deal with size frequency distribution (collisional cascade) resultant from collisions?

	We assume that there is a direct correspondence between the local collision rate and the amount of dust present. As the referee points out below, the spatial evolution of bodies between dust and planetesimal sizes might have a significant influence of the final dust distribution. We have made sure to highlight this concern throughout the paper, and provide a somewhat crude calculation in section 5 which suggests that this assumption is not terribly unreasonable.

1.4) What is the dust size considered in this work?

	We assume that the resulting dust grains are 1mm in size. This is mentioned in section 5 (section 4.2 in the original draft).

I would say all these matter a lot once 300 km planetesimals will likely not be instantaneously pulverized to dust-size when they collide. Thus it is of major importance for the understanding of this work, to know precisely how collisions were treated. If, for instance, authors did not consider any fragmentation, and only tracked collisions to determine where dust would form, I would say this should be put as a caveat.

The reason why I'm asking this to be a caveat is because in this case the authors' results would not necessarily be representing the dust profile, but rather the planetesimal collisional profile, and so the upper limit of dust generation at a given location. This makes a difference regarding the applicability of this work, once the expected real signals from the bumps and dips would be much smaller, and so particularly undetectable. The conclusion's by Wallace et al., potentially based on a local upper limit of dust generation, already points towards an expected faint signal in real observation. Therefore, if the real expected signal end up being even fainter, would this work results be applicable at all, or would them be simply negative results? I am afraid that if the results would be simply negative, they may not be important. I mean, why should we care if we will not be able to easily see those bump/dip features? In this case we would not be able to say anything about the mass and eccentricity of the potential perturbing planet. Therefore, what is the meaning of constraining something that it is unmeasurable?

I would try to help the authors and turn this into a positive light, even if the results turn to be negative. One could eventually say that, once observational capabilities improve, having a constrained measurement would help to quickly interpret images in the future. I don't know how likely is this, but with a proper caveat, it would be acceptable, eventually.

	This is an excellent point brought up by the referee and is something that deserves a more careful discussion of in the paper. In regards to the applicability of this work: one must first understand where and how the planetsimals collide if there is any hope of nailing down what the resulting dust distribution will look like. Going from the planetesimal collision profile to a dust emission profile deserves an entire paper by itself and we certainly do not do that analysis justice here.

	In light of this, we have decided that the best thing to do is to reframe this paper a bit, so that the collisions->dust step is saved until the end, given a little less weight in the abstract, introduction and conclusion, and the concerns brought up by the referee are sufficiently emphasized in the text. Most importantly, the reader is now cautioned that the dust profiles we present are upper limits.

	Both the abstract and the summary at the end of the introduction have been reworked to place more emphasis on the planetesimal dynamics and collisions, with the dust profiles and giant planet property constraints as more of an afterthought. Sections 2, 3 and 4 of the paper now focus solely on the full-sized planetesimals. All discussion of dust production and its observability is now saved for the latter parts of the paper.

2) I am very concerned by the fact the authors neglected migration on small particles due to aerodynamic drag. 300 km planetesimals will not break down to dust directly (of course some dust will be created during the 300 km planetesimal's collisions, but that would be a very small fraction of mass). The 300 km planetesimals (depending on material) will more likely break into a collection of smaller km-size objects. These smaller size objects will then overpopulate the regions where they were created, thus enhancing collision rates among them. Collisions that continue to happen among these smaller objets will eventually turn them in to a very large swarm of even smaller m- to km-size objects. Ultimately, this last generation of collisionally evolved planetesimals will become dust (then my question about how and if Wallace et al. considered collisional cascade in their simulations).

The problem of neglecting aerodynamic drag becomes then a bigger problem because, as it is very well known, small m- to km-sized objects feel strong headwind and thus can drift very fast Sunwards (e.g. Weidenschilling & Davis 1985; Patterson 1987; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995a,b; Jiang & Yeh 2004; Batigyn & Laughlin 2015; Storch & Batygin 2019; Deienno et al. 2020). Therefore, dust created by these large moving swarms could leave a signature that is far different than that when assuming dust would be generated in the location of the 300 km planetesimal collisions (eventually changing, or even destroying, the bump/dip signatures; despite the fact that dust would couple to the gas almost instantaneously, once created). This is why I previously said authors may be tracking the planetesimal collisional profile and the upper limit of dust generation at those locations in their simulations, but not necessarily the dust profile.

	In the original paper, we checked to ensure that the 300 km planetsimals and the mm-sized (and smaller) dust grains would stay in place, but the referee's concern about the effects of radial drift of intermediate-sized bodies is an important point that was overlooked. We handle this by: 1. cautioning the reader that the dust emission profiles presented should be interpreted as upper limits 2. discussing the concerns brought up by the referee 3. Providing a (somewhat crude) calculation of the drift and collision timescale of 1m bodies, where drift should be maximal. Point 3 is discussed more below.

I would suggest that the authors investigate and say explicitly in the text what is the expected relation between collisional timescale and drift timescale for m- to km-sized (1 m - 1 km) particles.

They should discuss:
2.1) If drift timescale of 1 m- to 1 km-sized objects is larger than their collision timescale.
In this case, collisions would happen before planetesimals would drift. Then the results should hold true.
2.2) If drift timescale of 1 m- to 1 km-sized objects is smaller than their collision timescale.
In this case, planetesimals would tend to drift faster than they can collisionally evolve. However, if their drift timescale is still larger than 5000 yr (integration time), then maybe this could be a minor issue. But a caveat would still be necessary as the large swarm of small planetesimals would likely continue collisionally evolving elsewhere (far from the mean motion resonances; MMRs, thus potentially changing the bump/dip signatures).
2.3) If drift timescale is smaller than 5000 yr and also smaller than collisional timescale.
In this case, this is a major issue! Planetesimals would quickly fall into the star, leaving minor dust signature, or a continue dust signal throughout their migration path, thus certainly no bumps nor dips would be observed.

Regardless the situation, this should be clearly discussed in the text.

	In section 5, we now provide a calculation of both the drift timescale and collision timescale for 1m bodies near the 2:1 MMR. Because we do not model the debris from planetesimal collisions, the collision timescale for the 1m bodies is found by extrapolating the collision rate of the planetesimals down to smaller bodies. This is done in the following way:

		1. Measure the collision rate (per particle) of the planetesimals from the e2m2 simulation inside the 2:1 MMR
		2. Assume that the 1m bodies will have the same dynamical state (eccentricity and inclination dispersion) as the planetesimals
		3. Assume that all of the mass in solids near the 2:1 MMR now consists of 1m bodies
		4. Scale the planetesimal collision rate (measured in the simulation) by the factor by which the number density and collision cross section would change if one were to convert the planetesimal population to smaller bodies. In this way, we get and estimate for the collision timescale for 1m bodies.

	In doing so, we estimate that the collision timescale for 1m bodies is up to 100x-1000x smaller than the drift time for those bodies across the resonance. This suggests that radial drift will not have a large influence on the resulting dust dustribution. However, the 1m collision timescale should be interpreted as an lower limit, due to assumptions 2. and 3 listed above. We conclude that it is *plausible* that radial drift does not play an important role, although a full, proper treatment of the collisional cascade and drift process is necessary to say this with certainty.

3) Wallace et al. mostly studied collisions that happen around MMRs. Thus, my point 2) above may become a problem for this as well. The pioneering work by Weidenschilling & Davis (1985) suggested small drifting particles could get captured by external MMRs with large planets. However, by that time the authors already pointed out that collisions would likely knock particles out of MMRs. More recently, Deienno et al. (2020) were finally able to confirm the predictions by Weidenschilling & Davis (1985) that collisions prevent small m- to km-sized particles to remain in MMRs (or near them) when gas is still around in the protoplanetary disk. Thus, I would say that, because Wallace et al. is mostly studying collisions within and near MMRs, in a protoplanetary disk (when gas should be around), adding aerodynamic drag on small size particles can be of a major influence, and eventually modifying all conclusions, specially because interior MMRs (those studied in this work) are largely inefficient in capturing any drifting small planetesimal (Weidenschilling & Davis 1985; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995a,b).

Summarizing, if interior MMRs can't capture small drifting planetesimals (or even in the case where capture would eventually occur), aerodynamic drag plus collisional evolution would not leave any object in MMRs. Therefore, the bumps and dips observed in this work would eventually not exist.

Not that I want to sound like a broken record, but in my opinion, without considering small planetesimals drifting due to gas drag, the figures presented in this work represent a planetesimal collisional profile, but not necessarily a dust profile. Again, large amounts of dust will not be directly generated by collisions among 300 km planetesimals, but rather by collisions among a larger swarm of m- to km-sized objects. The latter are strongly affected by the gas and don't strongly interact with interior MMRs, unless the planet in migrating Sunwards and capturing them (not the case of the present work).

    See our response to concern 2). The referee is absolutely correct that the connection we have made between the planetesimal collision profile and the dust profile is rather tenuous.

4) Wallace et al. performed each simulation for 5000 years. In the introduction, authors claim all works to date are severely limited in simulation time. Well, is 5000 years different than that? Deienno et al. (2020) performed N-body simulations of collisionaly evolving planetesimals (using statistical routines and tracer particles) with N of order O^10+ for 100,000 yr! I understand the codes used and the purpose of the two works were very different, so as the simulation setups. However, I don't understand why 5000 yr is not a severely limited time either. Protoplanetary disks can live much longer than that (Myr timescales). Why 400 orbits is enough? Wouldn't a longer simulation time change all results? I mean, the way it is, I understand that the disk signatures would be those reflecting a 5000 yr evolution, which is basically the observational instant considering how long a disk could live. However, if a 10,000 yr or 100,000 yr evolution would change the dust signatures (especially if considering aerodynamic drag), how applicable are the conclusions in this work? See that 10,000 yr or 100,000 yr are also short periods compared to the disk lifetime. Could you, please, add discussion regarding simulation time, as well as, clearly define what is a short and a long time (and why)? For me 5000 yr is, in principle, too short.

	Other studies which directly resolve planetesimal collisions near MMR with N-body methods (Richarson et al 2000, Dobinson 2013) used an integration time that shorter than the libration timescale of the resonances. This is what we mean by 'severely limited'. The text in the introduction has been changed to clarify this.

	Regarding other codes that make use of tracer particles (such as Deienno et al. 2020), one would worry that the phase space resolution may not be sufficient to fully resolve the dynamics that are happening, particularly near the edges of the resonances. It is doubtful that statistical routines would capture this either, unless it was explicitly built in to the code.

	Given that gas drag does not have a significant influence on the planetsimals (and it's plausible that the same is true for the collisional debris), we do not expect that a longer integration time should affect our results. The integration time we have chosen (5,000 years) is longer than any of the relevant timescales relating to the dynamics of the planetsimals. We now present a calculation all of the relevant timescales relating to the dynamics of the planetesimals in section 2. We have also added an expanded discussion justifying our choice of integration time to the beginning of section 4. In addition, section 5 contains a discussion of timescales relating to the debris from the collisional cascade process.

I'm very concerned about all these 4 points and would like that the authors pay careful attention, as well as, devote a thorough discuss related in the revised text.

The question of applicability:

If I understood the applicability of this work would follow the steps:
1) Once a bump or dip is observed in the dust emission signature of a disk, one would need to assume that this bump or dip reflects the position of the interior 2:1 MMR with a planet.
2) Thus, one would need to look for another dip or bump near what would be the interior 3:1 MMR with that supposed planet.
3) Only then one would be able to guess on the eccentricity and mass of the assumed perturber.

I think this should be clearly said in the conclusions, as well as this idea should be introduced somewhere upfront in the manuscript (maybe in the abstract or introduction). I would not leave to the reader to reach this conclusion. Readers can easily interpret things wrong (especially nowadays where we have so much things to read and we usually read them fast). Did I have the right interpretation or are the author suggesting something different in regards to applicability?

    This is the correct interpretation. The outline at the end of the introduction now lays this out. In addition, there is a step-by-step description of this process toward the end of section 7 (section 6 in the original draft).

General points (not in order of importance, but as they appear in the text):

I will use the following nomenclature:
- page 2, left(right) column, third paragraph, line 36 as pg.2 l(r)col. P.3 l.36 (line number reflects the number on the left side of a given page and not the paragraph line.)

1) pg.1 lcol P.1 l.47 - "or orbiting externally" - why not considering the planets orbiting internally too? External MMRs are important and can have a much larger effect in debris disk.

    This is absolutely true. The initial conditions for this work are meant to be compared with simulation B of Richardson et al. 2000, which only focused on the internal MMRs. The phrase 'orbiting externally' was changed to 'orbiting adjacent to'. The reason we did not explore exterior MMRs in this work is because doing so is more computationally expensive. (Many more orbits required for collisions to occur)

2) pg.1 lcol P.1 l.49-51 - "In gas-poor/debris systems, a giant planet perturber can also influence the structure of the dust continuum emission." - Why discussing debris disk if you will only account for the gaseous phase? This confuses the reader.

	Although the presence of gas is necessary in order to justify a simple, tractable model for connecting the planetesimal collision profile to the dust profile, the dynamics of the planetesimals and the resulting planetesimal collision profile does not depend very strongly on the gas. In a debris disk, we can't connect the planetesimal collision profile and the dust profile in such a simple way, but the behavior of the planetesimals should be quite similar.

3) pg.1 rcol P.2 l.44-47 - "Dobinson et al. (2013) showed that collisional dust that... to infer the presence of the planet." - In a debris disk or in a gas disk?

    The Dobinson et al. 2013 work modeled the inner, gas-poor region of a transition disk. The term 'planetesimal disk' in our introdution was changed to 'inner part of a transition disk' to make this more clear.

4) pg.2 lcol P.1 l.44-47 - "Unfortunately, collision detection in an N-body simulation is extremely computationally expensive. So far, studies of planetesimal dynamics near MMRs have involved either collisionless test particles (Boley 2017; Tabeshian &Wiegert 2016, 2018) or severely limited integration times (Richardson et al. 2000). To further elucidate this subject, we use the tree based N-body" - But see Deienno et al. (2020).

    Deienno et al. 2020 makes use of tracer particles, which is not the case with the other studies we cite here. Had we used coarser resolution and taken advantage of tracer particles in this work, the structure seen at the edges of the resonances in figures 7 and 8 (figures 6 and 7 in the original paper), which is critical to generating the features seen in the collision profiles, is unlikely to have been resolved. The text in the above sentence in the introduction has been modified to clarify that we are only referring to cases where the planetesimals are directly modeled.

5) A general comment - The introduction is confusing. It is clear from the abstract that the work will consider only the gas disk phase of a protoplanetary disk. However, because the introduction discuss a lot of debris disk, it is not clear whether debris disks will be considered. I would suggest focusing the discussion on protoplanetary disk. This way it would be clear authors will work during the gas phase.

	We have made edits to the abstract, introduction and conclusion to place more emphasis on the behavior of the planetesimals, which should be similar in both a protoplanetary disk and a debris disk. As we show in section 2.4 (newly added), the gas has an incredibly weak effect on the planetesimals and their resulting motion and collisions should be similar in both a gas-rich and gas-poor system.

6) Another general comment, now regarding section 2. - A much more relevant dynamics not discussed at all is due to aerodynamic gas-drag on planetesimals (especially the smaller ones). Also, collision rates are not discussed at all !

   Good point! Two additional subsections have been added to section 2, which discuss both gas drag on planetsimals and collisions between planetesimals.

7) Reference - Murray & Dermott (2000) - I think Murray & Dermott (1999) is a better reference (at least more commonly used).
\bibitem[\protect\citeauthoryear{Murray \& Dermott}{1999}]{1999ssd..book.....M} Murray C.~D., Dermott S.~F., 1999, ssd..book

    Never noticed this before! All Murray & Dermott references have been changed to the 1999 version.

8) pg.2 rcol P.1 l.8-9 - "... along with aerodynamic gas drag will..." - Even at this point is still not clear from the text if gas will be considered. Please, clarify the text.3

    This is now clarified at the end of section 2.4 (gas drag on planetesimals) and in section 3.2 (initial conditions). The idea is that we consider the effects gas drag when building the initial conditions, but do not need directly model it during the course of the simulations because the gas drag and gravitational stirring timescales (which counteract each other) are quite long. We assume that these two effects are in equilibrium, build them into the initial conditions (through the eccentricity and inclination dispersions) and then can safely forget about them.

9) A general comment with respect to section 2.2 - It looks like there are typos in some of the equations. I would also suggest that, every time that equations from other works are quoted, please, say chapter, page, and eq #. This helps the reader to look after, if a bit more of background is needed. Finally, I don't understand the importance of this section. It ends as: "Due to the complicated dynamics that are at play in this region, we turn to a direct N-body treatment to better understand how the collision rate varies in the vicinity of a MMR.". When I reached this point I asked myself: Will this section be used for something or is it here simply to say that the dynamics are complicated enough to be treated analytically, thus needing numerical analysis? I would suggest to shorten this section. Another suggestion, consider moving it to an appendix (once equations are revised).

	Chapter, page number and equation number have now been added alongside all equations pulled from Murray and Dermott. This section has been repurposed a bit, and some of the more detailed parts have been moved to the appendix. The main purpose of this section is now to present calculations for the timescales of dynamical effects (secular forcing, libration in MMR, aerodynamic gas drag) that act on the full-sized planetesimals. Note that all of the discussion and calculations of the dynamics of the collisional debris is saved for section 5. Most of the detailed references to Murray and Dermott have been moved to the appendix, as this is probably an unnecessary amount of detail for the typical reader.

I am not saying equations are wrong, but please, revise all equations to make sure they are correct and consistent, and add explanation when clarification is needed.

10) Equation (5) - should'n this equation have -q instead of -(2p+q) in the third term and -p instead of -q in the second term? See for example Murray & Dermott (1999, chap 8, eq. 8.8), or Beaugé & Ferraz-Mello (1993), or Malhotra (1993), or Gomes (1995)...
From Murray & Dermott (1999, Chap 6 and 8), one of the general forms of the resonant angle can be phi = j1*lamb' + j2*lamb + j4*varpi, where j1 = p+q, j2 = -p and j4 = -q. So why using j2 = -q and j4 = -(2p+q) instead?
The authors should explicitly say, and explain why these choices were made.

    The argument should indeed by -q, rather than -(2p+q). This mistake was carried through when writing up the equations in the paper, although it appears we properly used j4 = -q when doing our analysis and constructing the plots. (2p+q) has been replaced by q in equations 5 and 6 in the original draft (equations 5 and B1 in the revised version).

11) Equation (6) - Assuming j4 = 2p+q

    See 10

12) pg.3 lcol P.1 l.11 - "... table 8.1)..." - third row, with j = p+q (in Murray's book).

    Added

13) Equation (9) - Shouldn't this be fd(alfa) = 1/2 [ -2j -alphaD ] b1/2^j (minus, not plus)

    You are correct. There was a minus sign missing in front of both terms in the paper. This has now been fixed.

14) pg.3 lcol P.2 l.21 - "... table 8.1)..." - first row in Murray's book

    Added

15) pg.3 lcol P.2 l.28 - "... equation 8.76)." - with j2 = p.

    Added

16) pg.3 lcol P.4 l.34-35 - "The net result of this is that MMRs will produce spikes in the semimajor axis-eccentricity distribution of a planetesimal disc." - This is true when migration due to aerodynamic drag is not considered, or for large planetesimals that don't feel strong headwind. Would this still be true for small planetesimals that are drifting inwards? That would largely depend on the relation strength of the resonance vs drifting timescale (resonance crossing time).

    This section is meant to focus solely on the dynamical effects that the planet has on the full-sized planetesimal population, in order to motivate the initial conditions used in the next section, and also to explain the planetesimal collision profiles shown in section 4.1. We are purposefully trying to keep discussion of radial drift and collisional debris as a separate, encapsulated part of the paper (section 5) in the new version.

17) A general note regarding section 3.1 - The code should be described with more details. This is a tree code primarily developed to work on cosmology simulations, thus not commonly applied for astronomical models. What is need to translate a cosmology code to astronomy? What are the differences in potential? I have many questions and a general reader would have them too. In my opinion, simply saying the code works well in half million processors and that it calculates gravitational forces, then redirecting to Jetley et al. (2008) is too superficial. See I am not asking to give all details, as this is an already published code. But having a better description of the portions of the code that are relevant to this work would be important. Also, I think it is crucial that Wallace et al. describe with more details how the code deals with collisions, and all the assumptions made, as well as all the consequences of it. Again, I understand the authors have a full description of the collisional module in Wallace & Quinn (2019), but having a summary of the main relevant features here would not hurt. On the opposite, it would greatly help the reader and the understanding of the work. Besides, including more details would not be a big task, and should also not compromise the length of the work (especially if Sect 2.2 goes mostly to an appendix). Please, I beg for more detail in this section.

    We provide some references to planet formation simulations done with PKDGRAV, which uses similar methods to ChaNGa. An additional paragraph has been added which summarizes some of the details provided about the ChaNGa collision model described in Wallace & Quinn 2019.

18) pg.3 rcol P.2 l.29-32 - "In the other four cases, the mass (eccentricity) is increased or decreased by a factor of two, while the eccentricity (mass) is held at the nominal value." - This reads awkward. Why not simply say that in other cases mass and/or eccentricity is increased or decreased by a factor of 2?

    Agreed, the referee's suggestion is much clearer.

19) pg.3 rcol P.2 l.33 - "... 5.2 AU orbit..." - 1) Why not closer distances? 2) AU -> au (here and elsewhere).

    Another sentence was added to the first paragraph of section 3.2 which explains our choice of parameters more clearly. The 'nominal' simulation is meant to be compared directly with simulation B of Richardson et al. 2000, although we reduce the width of the planetesimal annulus to reduce the number of particles required (while still covering all of the promininent interior MMRs). AU has been changed to au everywhere in the paper.

20) pg.3 rcol P.2 l.34 - "The planetesimal disc extends from 2.2 to 3.8 AU..." - Why not consider the outer disk too? External resonances can be very strong and capture is more likely there.

    This is a good point. Our work was mainly motivated by the non-result seen by Richardson et al. 2000, which is why we focus on interior MMRs here. We have also added a brief discussion highlighting the importance of studying exterior resonances in future work in the summary and discussion near the end of the paper.

21) pg.3 rcol P.2 l.41 - This is a minor comment. I am just curious. Why not using alpha = -1 or -0.9 as it seems more common for protoplanetary disks? -1.5 is an approximation for the Minimum Mass Solar Nebula based on the current solar system planets. Observations of proptoplanetary disks sugggest alpha = -1 or -0.9.

    See 19) and 20). alpha = -1,5 has been a standard benchmark in the field for quite a while and is we do not think it's unreasonable to use. Furthermore, there is evidence to suggest that the solid surface density profile close to the star is steeper than alpha = -1 (see Chiang + Laughlin 2013, and Dai, F 2020). The region of the protoplanetary disk that we are modeling is not well-constrained by either exoplanet studies or mm and sub-mm observations.

22) pg.3 rcol P.2 l.44 - "Each simulation is evolved for 5,000 years" - Could you, please, add discussion regarding simulation time, as well as, define what is a short and long time. For me 5000 yr is, in principle, too short (see my concern #4).

    See our response to concern #4. Some text has been added to the paper (at the beginning of section 4) to elaborate on this.

23) pg.3 rcol P.3 l.44-46 - "The dynamical effects of secular forcing by the giant planet, along with the effects of viscous stirring and gas drag on the planetesimals, are built into the initial conditions." - From the text, It is not clear whether the authors are considering the gas disk phase or not?

    See 8). We are considering the gas disk phase here. This shouldn't matter too much in terms of the motion and collisions of the planetesimals, although the presence of the gas is very important to construct the dust profiles. More text has been added to this section to explain this more clearly.

24) pg.3 rcol P.3 l.48 - "... and gas drag as a function of semimajor axis..." - Again, is this only used to setup the disk condition, or is this updated during the simulation, once debris are created? Are debris created?

    See 23). We do not keep track of debris during the simulation. A note has been added to the end of section 3.1 which states this.

25) pg.3 rcol P.3 l.49-50 - "Although ChaNGa does not account for the effects of gas drag on the planetesimals..." - Please, see my concerns and add discussion.

    See 23) and 24)

26) pg.4 lcol P.2 l.49 - "dynamical timescale of planetesimals..." - What do you mean by dynamical timescale of planetesimals? Collisional timescale, drift timescale, what?

    This is the timescale for gravitational interactions between planetesimals that undergo close encounters. A sentence has been added to this section to clarify.

27) pg.4 rcol P.1 l.41 - "... T_lib ~ 1,000 - 2,000 years..." - So the integration time is about a factor of 2-4 T_lib. Is this enough to take conclusions? Please, see my concern and add discussion.

    In our case, the large particle count allows us to look at the 'ensemble average' of the dynamical behavior near resonance. Even though individual planetesimals only go through a few librational oscillations, the phase trajectories that they follow are well populated with particles. This allows us to make conclusions about the dynamics without needing a longer baseline from which to take time averages. Had we been following only a few bodies, this would not have been the case. Some additional text has been added to this paragraph to explain this more clearly.

28) pg.4 rcol P.2 l.52-53 - "Resonances with the perturbing planet are visible as nonaxisymmetric gaps." - I am afraid that this is only because none of your particles are drifting, which would locally increase the collision rates once small planetesimals fragments that should be drifting away start to (maybe unrealistically) pile up within or near MMR, thus quickly collisionally evolving to dust. I wonder if this would ever happen if gas drag was to play? Please, see my concern and add discussion.

    This is referring to the positions of the planetesimals in this figure, and not to any kind of debris. Even if the smaller fragments do drift significantly, the nonaxisymmetric gaps in the planetesimal distribution should still be present. Of course, this would wash out all of the interesting features in the dust emission profile. This sentence has been modified to clarify that we are specifically talking about the structure seen in figure 1.

29) pg.4 rcol P.2 l.56-59 - "... in Richardson et al. (2000) (see bottom panel of figure 3) and Tabeshian & Wiegert (2016) (figure 3)." - add figure 3 in those works. Fig 3 in this work does not have bottom panels.

    The text now specifies the subfigures we are referring to (3c and 3f in Richarson, 3a in Tabeshian & Weigert)

30) Figure 2 - I wonder how (or if) particle drifting due to gas-drag would change this figure? Also, the 8:3 and 7:2 MMRs are irrelevant here! Why not showing 7:3 MMR that is discussed in the main text?

    As we now calculate in section 2.1, the libration timescale and the secular forcing timescale are much shorter than the stopping time due to gas for the planetesimals, so this figure should not look different had we modeled gas drag in the simulations. A typo in the text has been corrected: '7:3' was supposed to say '8:3'.

31) Caption of Figure 2 - "Between the resonances, the nonzero eccentricity is due to secular forcing by the planet." - Gas would likely damp them to zero!

    See 30)

32) pg.5 rcol P.1 l.15 - "PDF" - Please, define abbreviations before using it.

    PDF stands for probability density function. This is now stated in the text.

33) pg.5 rcol P.2 l.24-27 - "The most prominent features appear as an enhancement to the collision rate near the 2:1 MMR and a decrease near the 3:1 MMR." - I don't see a decrease near 3:1 in figure 4. I see two peaks around 3:1 and one interior to the 2:1.

    Perhaps we should have been more careful with language here. In figure 4, there is a local minimum near the 3:1 MMR. There are certainly some additional features surrounding it, although these are partly due to collisions with bodies on high eccentricity orbits in the adjacent (7:2 and 5:2) MMRs. The text has been changed to refer to the features at the 3:1 and 2:1 MMRs as local maxima and minima.

34) Last paragraph of section 4.1 - "We attribute... radial dust distribution" - I don't understand this argument about gravitational focusing. In my interpretation, what could be happening is that planetesimals that get excited by the MMRs collide at their perihelion distances. This explains why the larger relative # of collisions are close to the MMRs and interior to it.

    An additional figure (fig 5 in the new paper) has been added to clarify the gravitational focusing explanation. The basic idea is that both a dynamically cold and a very dynamically hot planetesimal population will exhibit a large collision rate. When things are cold, gravitational focusing is highly effective. When things are very hot, encounters are very frequent and many collsions occur, even though the effective collision cross section is diminished. The collision rate exhibits a minimum value at some intermediate value of dynamical excitation. Depending on how dynamically hot the local planetesimal population gets when a resonance heats it up, the collision rate can either increase OR decrease from the unperturbed value.

    To address the referee's intrepretation: if the offset in the peak of the collision rate in figure 4 (near the 2:1 MMR, for example) were due to bodies colliding at perihelion, the peak should shift inward as the eccentricity of the planet is increased. Between the e1m2 and e3m2 simulations, the mean eccentricity at the 2:1 MMR increases from ~0.03 to ~0.06 and so the perihelion distance of a body at the 2:1 MMR would decrease by about 0.1 AU. This shift would have apparent in figure 4.

35) pg.6 lcol P.1 l.33 - "cylindrical distance" - please, define before start using it.

    A sentence has been added to the text to clarify what cylindrical distance means.

36) pg.6 rcol P.2 l.47-54 - "This also suggests... radial collision profile." - I don't know if I agree with authors' conclusions that MMRs are not that important, and that most collisions occur with objects out of resonance. It looks for me that the features seem in Fig 6 (similar to Fig 4) are caused by objects in MMRs colliding at their pericentric and apocentric distances within the eccentric orbits. In my understanding, if authors' arguments are correct, MMRs should basically not matter, and Fig 5 would look similar if no planet was present. If that is the case, how can this work serve to constrain the mass and eccentricity of a perturbing planet?

    As shown in figure 6 (fig 5 in the original version), the bump and dip features at the center of the MMRs are still present, even if we ignore particles whose semimajor axis falls within the libration width of the resonances. What this implies is that the edges of the resonances are important. As we approach a resonance from the outside, the critical angle \phi becomes near-stationary (using the pendulum analogy, this corresponds to a bob in the vertical, upside-down position). Where the critical angle is near-stationary, the pericenter orientations of the planetesimals experience a torque which causes their \varpi to follow the giant planet in its orbit. This is what we propose is responsible for the coherent structure seen in the right-hand panels of figures 7 and 8 (6 and 7 in the original paper), along with the structure seen in the collision profiles.

    The paragraph referenced now clarifies that we are specifically referring to collisions between bodies inside of the libration width of the resonance. An additional sentence has been added at the end of the following paragraph (which provides an explanation for the nonaxisymmetric gaps seen in fig 1) to tie this back to figure 6.

37) pg.6 rcol P.2 l.38-56 - "As mentioned in section 2, ... appears to 'follow' the planet." - This paragraph is super confuse. I don't understand this. Why not simply say this is caused by the secular forcing in eccentricity? This would save 3 paragraphs in the paper and would be much more clear.

    See 36). If you look at figures 1, 7 and 8, (1, 6 and 7 in the original paper) the nonaxisymmetric gap structure in the disk near the resonances is always aligned with the planet's orbital phase. It is also apparent from figure 1 that the gaps are not aligned with the planet's longitude of pericenter, which should be the case if the structure were due to secular forcing in eccentricity. An explanation similar to the referee's was given by Tabeshian & Weigert 2016, which, as we point out, does not explain why the gaps are misaligned from the planet's longitude of pericenter.

38) Figs 4 and 5 - Why is the relative # larger than 1? Relative to what?

    These are showing the probability density of the collision rates as a function of semimajor axis (fig 4) and distance from the star (fig 5). The curves on these plots are normalized such that the area underneath = 1. For the orange curve in figure 5, the normalization factor is scaled by the ratio between the total number of collisions in each sample. Some text has been added alongside the discussion of figures 4 and 5 to explain this.

39) General note about the figures - Not that I want to be repetitive, but I am afraid that if small planetesimals are allowed to drift the bumps and dips from these figures may change as the small planetesimals would likely collide everywhere in radial distances. This is because internal MMRs are not strong enough to perturb these small fast-drifting planetesimals. In my opinion, without considering small planetesimals drifting due to gas drag, Figs 6 and 7 (especially) represent a planetesimal collisional profile, but not necessarily a dust profile. Please, see my concerns and add discussion.

    See our response to concern 4)

40) pg.10 lcol P.4 l.42-44 - "Many different parameterizations are possible, depending on optical depth, grain size, disc temperature, etc.,..." - But couldn't this change all of your conclusions? Please, see my concerns regarding planetesimal sizes and add discussion.

    To clarify, these parameters will alter the overall brightness profile of the dust, but should not affect the features due to the resonances. Some additional text has been added to this section to explain this.

41) Last paragraph of section 5 - "In order to... cases shown above" - This is a big IF, especially if what the authors are measuring here is the planetesimal collisional profile rather than the dust profile, or if the dust profile would be even shallower as a consequence of aerodynamic drag acting on small fragments of collisionally evolved planetesimals.

    This is certainly a big 'if', although we have shown with a back-of-the-envelope calculation in section 5 that the connection between the collision profile and the dust profile is not completely unreasonable. We have added some additional text to the end of section 5 to remind the reader that the dust profile is built on somewhat tenuous assumptions.

42) pg.12 rcol P.1 l.8-24 - "Near the interior 2:1... both of these quantities." - This is a clear way to set the explanation. Well done! I hope to see more of that in the revised version. I think a lot of the explanations in the main text should be substituted by something clear and concise as this! Make it simple!

    :-)

43) pg.12 rcol P.3 l.45-48 - "Given that the solid surface density diminishes with distance (r^-1.5 for the MMSN (Hayashi 1981)), exterior resonances may not be quite as useful for this type of measurement, however.)" - Not sure I agree. Exterior MMRs can more easily capture inward drifting objects (e.g. Weidenschilling & Davis 1985; Patterson 1987; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995; Jiang & Yeh 2004; Batigyn & Laughlin 2015; Storch & Batygin 2019), increasing their local collision rates (Deienno et al. 2020), thus potentially generating very strong dusty features.

    This is absolutely true if we are going to consider radial drift. Given that an expanded discussion of radial drift is now an important part of this paper, we should be more careful not to dismiss the importance of exterior MMRs. Here, we have added a sentence mentioning that exterior MMRs might be important if radial drift is modeled.

44) pg.12 rcol P.4 l.50-52 - "As we have shown, the bump vs dip feature near the 2:1 MMR is marginally detectable with ALMA for the nearest protoplanetary discs." - Please, see my concern and add discussion.

    This is addressed in a few other places in the new version of the paper, but we have added a sentence here reminding the reader of the assumption that goes into constucting the dust profile.

My apologies for being so long. I do think this work has potential, but it does need substantial clarifications.

	No need to apologize. Thank you very much for being so thorough!

	References used in response:

	- Chiang, E. & Laughlin, G., 2013, MNRAS, 431, 3444C
	- Dai, F. et al., 2020, AJ, 159, 247D